---
title: Data Viz
author: Chris Smith
date: '2017-12-12'
slug: data-viz
categories: []
tags: []
---

In R we can create cool data visualizations that represent data in an easy and communicative way.
First, we must qurey our data into dataframes with sqldf(). 
```{r}
library(Lahman)
library(sqldf)
library(RSQLite)
#Data visulization scatter plots
#extracting Data 
query <- "SELECT playerID, sum(HR) AS Career_HR, sum(SO) AS Career_SO
FROM Batting 
GROUP BY playerID 
HAVING sum(HR)>=400"



sqldf(query)

plotdf <- data.frame(sqldf(query))
```

We then can begin using ggplot to create our visualizations. There are three main componets to ggplot: data, geometery and aesthetics.
```{r}
library(ggplot2)
str(plotdf)


plot1 <- ggplot(data=plotdf, aes(x=Career_SO, y=Career_HR, label=playerID)) + 
  geom_point()+
  geom_text(aes(label=playerID),hjust=0, vjust=0)+
  ggtitle("Career Homeruns VS Career Strikeouts for Great Hitters")+
  xlab("Career Strikeouts")+
  ylab("Career Home Runs")
plot1
```

We also can make histrograms.
```{r}
#Histogram ggplot

#Extracting Data
query2 <- "SELECT weight FROM Master"
plotdf_2<-sqldf(query2)

#Data Viz

plot2 <- ggplot()+
  geom_histogram(data=plotdf_2, aes(x=weight), color="red", fill="white")+
  ggtitle("Ball Player Bodyweight Distribution")
plot2

```

We also can create times series analysis using ggplot as well.
```{r}
#Time Series Plot for the Great Bambino 

#Extracting Data
query3 <- "select yearID, HR from Batting
where playerID='ruthba01'"

plotdf_3<-sqldf(query3)

#Data Viz 

plot3 <- ggplot()+
  geom_line(data=plotdf_3, aes(x=yearID, y=HR),color='blue')+
  geom_point(data=plotdf_3, aes(x=yearID, y=HR),color='red')+
  ggtitle("The Sultan of Swat's Homerun Blasts Through the Years")+
  xlab("Year")+
  ylab("Homerun Blasts")
plot3
```

Using ggplot we can also create bar plots as well.
```{r}
#barplot
#extracting the data
colnames(Teams)

query4 <- 'select name, HR from Teams
where yearID=1980
order by HR'
plotdf_4 <- sqldf(query4)

#makeing the barplot 

plotdf_4$name<-factor(plotdf_4$name, levels=plotdf_4$name)


plot4 <- ggplot(data=plotdf_4, aes(x=name, y=HR, fill=HR))+
  geom_bar(stat='identity')+
  xlab('Homeruns')+
  ylab('Team Names')+
  ggtitle('1980 Team Homerun Distribution')+
  coord_flip()
  
plot4
```

